---
title: "fig3a_bray_curtis_plots"
author: "Minghao Chia"
date: "2/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(reshape2)
```

Read in data.

```{r}
metadata <- read_tsv("../metadata/sample_metadata_counts.tsv", col_names = TRUE) %>%
  filter(after_decont >= 10000)
load(file="../data/processed/subjects_pcoa_BC.Rdata") #subjects_pcoa_BC
```

Load functions

```{r}

###Remove mirror comparisons i.e reciprocal comparisons using a function, taking subjects_BC_pairwise
#The function takes subjects_BC_pairwise as an input. Set "Var1" as strictly for the children. 
#Set "Var2" as strictly for the caregivers. 

remove_mirror_comparisons_fn <- function(x){
  x <- filter(x, Var1_sample_cat == "child")
  list_by_Var1 <- split(x, f= x$Var1 )  #splitting removes mirror comparisons
  
  #Now select for pairs for which Var2_sample_cat is "caregiver"
  
  list_by_Var1_Var2 <- lapply(list_by_Var1, function(input){
    return(input %>% group_by(Var1) %>% subset(Var1_sample_cat != Var2_sample_cat)) 
  })
    
  output <- do.call("rbind", list_by_Var1_Var2)
  return(output)
}

#############################################################

#Avoid using rbind from base R when dealing with tibbles. Use dplyr::bind_rows() instead
#body_region can be either "VF", "AC", "LE" or "CH" and must be specified for both child and caregiver
#disease_state is either 0 (healthy) or 1 (AD)
#Input/x is the subjects_BC_no_mirrors dataframe

match_for_boxplot_fn <- function(x, child_body_region, caregiver_body_region, disease_state){
  child_body_region <- enquo(child_body_region)
  caregiver_body_region <- enquo(caregiver_body_region)
  disease_state <- enquo(disease_state)
  same_hse <- x %>% filter(Var1_region == !!child_body_region & Var2_region == !!caregiver_body_region) %>%
    filter(Var1_child_disease_state == !!disease_state, Var2_child_disease_state == !!disease_state) %>%
    filter(Var1_hse_grp == Var2_hse_grp)
  
  diff_hse <- all_discordant_pairs %>% 
    filter(Var1_region == !!child_body_region & Var2_region == !!caregiver_body_region) %>% 
    filter(Var1_child_disease_state == !!disease_state, Var2_child_disease_state == !!disease_state)
  
  same_hse$comparator <- ifelse(same_hse$Var1_child_disease_state == 0 &
                                           same_hse$Var2_child_disease_state == 0, 
                                           "healthy_matched_house", "AD_matched_house")
  
  
  diff_hse$comparator <- ifelse(diff_hse$Var1_child_disease_state == 0 &
                                           diff_hse$Var2_child_disease_state == 0, 
                                           "healthy_unmatched_house", "AD_unmatched_house")
    
  output <- dplyr::bind_rows(same_hse, diff_hse)  #do not use rbind()
  return(output)
}

####################################################################

#Get all discordant pairs (different hse_grp, different sample_cat (child vs caregiver)) per unique sample in "Var1". Note that "Var1" refers to children, "Var2 refers to caregivers
#The function takes subjects_BC_no_mirrors as an input. Value refers to the Bray Curtis distance between the pairs.
get_all_discordant_pairs_fn <- function(x){
  discordant <- x %>% group_by(Var1) %>% subset(Var1_hse_grp != Var2_hse_grp ) %>% 
    subset(Var1_sample_cat != Var2_sample_cat) 
  return(discordant)
}


```

Format the dist matrices and dataframes.

```{r}

#Melt the distance matrix of BC scores
# By dropping zero, we remove all self-comparisons
subjects_BC_pairwise <- subset(melt(as.matrix(subjects_pcoa_BC)), value!=0)

#Add meta info for subsetting (say healthy vs diseased households)

subjects_BC_pairwise <- metadata %>% subset(select = c(body_region, sample_category, hse_grp, subj_region_label, child_disease_state )) %>%
  rename(Var1_region = body_region, Var1_sample_cat = sample_category, Var1_hse_grp= hse_grp , Var1 = subj_region_label, Var1_child_disease_state = child_disease_state) %>%
  merge(subjects_BC_pairwise, by = "Var1")


subjects_BC_pairwise <- metadata %>% subset(select = c(body_region, sample_category, hse_grp, subj_region_label, child_disease_state )) %>%
  rename(Var2_region = body_region, Var2_sample_cat = sample_category, Var2_hse_grp= hse_grp , Var2 = subj_region_label, Var2_child_disease_state = child_disease_state) %>%
  merge(subjects_BC_pairwise, by = "Var2")

subjects_BC_no_mirrors <- remove_mirror_comparisons_fn(subjects_BC_pairwise)


all_discordant_pairs <- get_all_discordant_pairs_fn(subjects_BC_no_mirrors)

```


Make boxplots of of Bray-Curtis distances. One boxplot for matched child-caregivers from same households, another boxplot for discordant child-caregivers from different households. 

Compare volar forearm of children and parents in healthy and diseased households (VF, both matched and unmatched/discordant)

```{r volar_BC_boxplots, echo=FALSE}


healthy_VF_BC_df <- match_for_boxplot_fn(subjects_BC_no_mirrors, 
                                         child_body_region = "VF", 
                                         caregiver_body_region = "VF", 
                                         disease_state = 0)

AD_VF_BC_df <- match_for_boxplot_fn(subjects_BC_no_mirrors, 
                                         child_body_region = "VF", 
                                         caregiver_body_region = "VF", 
                                         disease_state = 1)

#Combine the above two dataframes
all_VF_BC_df <- rbind(healthy_VF_BC_df, AD_VF_BC_df)



my_comparisons <- list( c("healthy_unmatched_house", "healthy_matched_house"), 
                        c("AD_unmatched_house", "healthy_matched_house"),
                        c("AD_matched_house", "AD_unmatched_house"))

#Make the plot

VF_BC_boxplot <- ggplot(all_VF_BC_df, aes(x=comparator, y=value)) + 
  geom_violin(scale="width") +
  geom_boxplot(fill=c("red", "red", "blue", "blue")) + 
  theme_classic() + 
  stat_compare_means(paired = FALSE, method = "wilcox.test", comparisons = my_comparisons) +
  ylab("Bray-Curtis dissimilarity") +
  ggtitle("Volar forearm (children vs caregivers)")

VF_BC_boxplot

ggsave(filename = "../plots/fig3a_VF_BC_boxplot.svg" ,plot=VF_BC_boxplot, width=18, height=10)
```

Compare antecubital fossae of children and parents in healthy and diseased households (AC, both matched and unmatched/discordant).


```{r antecubital_BC_boxplots, echo = FALSE}

healthy_AC_BC_df <- match_for_boxplot_fn(subjects_BC_no_mirrors, 
                                         child_body_region = "AC", 
                                         caregiver_body_region = "AC", 
                                         disease_state = 0)

AD_AC_BC_df <- match_for_boxplot_fn(subjects_BC_no_mirrors, 
                                         child_body_region = "AC", 
                                         caregiver_body_region = "AC", 
                                         disease_state = 1)

all_AC_BC_df <- rbind(healthy_AC_BC_df, AD_AC_BC_df)

#Make the plot

AC_BC_boxplot <- ggplot(all_AC_BC_df, aes(x=comparator, y=value)) + 
  geom_violin(scale="width") +
  geom_boxplot(fill=c("red", "red", "blue", "blue")) + theme_classic() + 
  stat_compare_means(paired = FALSE, method = "wilcox.test", comparisons = my_comparisons) +
  ylab("Bray-Curtis dissimilarity") +
  ggtitle("Antecubital fossae (children vs caregivers)")

AC_BC_boxplot

ggsave(filename = "../plots/fig3a_AC_BC_boxplot.svg" ,plot=AC_BC_boxplot, width=18, height=10)

```
